Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-25 Paper/Paper.tex
minor
conclusion improved

We use an online version of the vector quantization algorithm
\cite{Gersho1991} to cluster the beat-chroma patches described in the
previous section.
% This can also be seen as an online K-means.
For each sample from the data, the algorithm finds the closest cluster
in the codebook and updates the cluster centroid (codeword) to be
closer to the sample according to a learning rate $\ell$.
The clusters are updated as each data point is seen, as opposed to
once per iteration in the standard $k$-means algorithm.
The details are explained in Algorithm \ref{algo:vq}.
%
As in standard $k$-means clustering, the codebook is initialized by
choosing $K$ random points from our dataset.
% ADDED BY TBM AFTER DAN
Initialization had no significant effect in our experiments.
%
Note that this algorithm, although not optimal, scales linearly with
the number of patches seen and can be interrupted at any time
to obtain an updated codebook. %not that this is important...
% The setting of the learning rate might become trickier, but this is
% not fundamentally a problem.

@@

The main future direction is larger scales. This implies $3$ improvements.
First, gather more songs from the Echo Nest API. Second, efficiently work
with larger codebooks. Some improvements have been published for $k$-means
that avoid comparing every code to every sample. Third, the learning
algorithm must not converge too fast. We could add a filter that selects
increasingly hard example to encode, our merge many codebooks in a similar
fashion as hierarchical Gaussian mixtures \cite{Vasconcelos2001}.

>>>>

We use an online version of the vector quantization algorithm
\cite{Gersho1991} to cluster the beat-chroma patches described in the
previous section.
% This can also be seen as an online K-means.
For each sample from the data, the algorithm finds the closest cluster
in the codebook and updates the cluster centroid (codeword) to be
closer to the sample according to a learning rate $\ell$.
The clusters are updated as each data point is seen, as opposed to
once per iteration in the standard $k$-means algorithm.
The details are explained in Algorithm \ref{algo:vq}.
%
As in standard $k$-means clustering, the codebook is initialized by
choosing $K$ random points from our dataset.
% ADDED BY TBM AFTER DAN
Initialization had no significant effect on the distortion
in our experiments.
%
Note that this algorithm, although not optimal, scales linearly with
the number of patches seen and can be interrupted at any time
to obtain an updated codebook. %not that this is important...
% The setting of the learning rate might become trickier, but this is
% not fundamentally a problem.

@@

The main future direction is larger scales. This implies $3$ improvements.
First, gather more songs from the Echo Nest API. Second, efficiently work
with larger codebooks. Some improvements have been published for $k$-means
that avoid comparing every code to every sample. Initialization
should also be modified, so every codeword is used (unlike in
Figure \ref{fig:freqs}). Third, the learning
algorithm must not converge too fast. We could add a filter that selects
increasingly hard example to encode, or merge many codebooks in a similar
fashion as hierarchical Gaussian mixtures \cite{Vasconcelos2001}.
